---
title: "Lab 11"
author: "Adam Hayes, Erin Omyer, Richard Park, Jasmine Sanchez"
date: "4/3/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(babynames)
```

###Ariel and Rachel Regexs

```{r}
Anames <- filter(babynames,str_detect(babynames$name,"A"),year == 1988,sex == "F")
Anames2 <- filter(babynames,str_detect(babynames$name,"A"),year==1990,sex=="F")

names <- inner_join(Anames,Anames2,by="name") %>% mutate(diff=prop.y-prop.x) %>% select(name,prop.x,prop.y,diff)

year1973 <- filter(babynames,str_detect(babynames$name,"Ra[ey]?ch.*l"),year == 1973, sex == "F") %>% count(wt=n) 
#8346

year1988 <- filter(babynames,str_detect(babynames$name,"Ra[ey]?ch.*l"),year == 1988, sex == "F") %>% count(wt=n) 
#19946

year1990 <- filter(babynames,str_detect(babynames$name,"Ra[ey]?ch.*l"),year == 1990,sex == "F") %>% count(wt=n) 
#20369

year2017 <- filter(babynames,str_detect(babynames$name,"Ra[ey]?ch.*l"),year == 2017,sex == "F") %>% count(wt=n) 

year1973_2 <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1973, sex == "F") %>% count(wt=n) 
#52

year1988_2 <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1988, sex == "F") %>% count(wt=n) 
#1617	

year1990_2 <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1990,sex == "F") %>% count(wt=n) 
#5366

year2017_2 <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 2017,sex == "F") %>% count(wt=n) 

Arieldiff <- filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1990,sex=="F") %>% count(wt=prop) - filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year==1988,sex=="F") %>% count(wt=prop)

Rachel <- filter(babynames,str_detect(babynames$name,"Ra[ey]?ch.*l"),year%in%c(1973,1988,1990)) %>% filter(name!="Rachelann") %>% filter(name!="Rachelanne")

Racheldiff <- filter(Rachel,year==1990,sex=="F") %>% count(wt=prop) - filter(Rachel,year==1988,sex=="F") %>% count(wt=prop)

```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1973, caption = "The number of females with the name 'Rachel' in 1973")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1988, caption = "The number of females with the name 'Rachel' in 1988")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1990, caption = "The number of females with the name 'Rachel' in 1990")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year2017, caption = "The number of females with the name 'Rachel' in 2017")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1973_2, caption = "The number of females with the name 'Ariel' in 1973")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1988_2, caption = "The number of females with the name 'Ariel' in 1988")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year1990_2, caption = "The number of females with the name 'Ariel' in 1990")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(year2017_2, caption = "The number of females with the name 'Ariel' in 2017")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(Arieldiff, caption = "The difference in proportion of the female Ariel baby names over time")
```
```{r echo = FALSE, results = 'asis'}
library(knitr)
kable(Racheldiff, caption = "The difference in proportion of the female Ariel baby names over time")
```
####Comments
Over the time between 1973 and 1990, the popularity of the name was overall more common than the name 'Ariel'. The name 'Ariel' spiked between the years 1989 and 1990 from 1,617 to 5,366 females with the name. The chance of the name 'Rachel' was more common than the name 'Ariel', due to the simplicity in the name 'Rachel'. In 2017, the name 'Rachel' has almost double number of females than the name 'Ariel'.

###The Little Mermaid Effect
```{r}
library(tidyverse)
library(babynames)

filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1973, sex == "F") %>% count(wt=prop) 
#3.346e-05

filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1988, sex == "F") %>% count(wt=prop) #0.000841196	

filter(babynames,str_detect(babynames$name,"Ar[iy]+.l+[^a]?$"),year == 1990,sex == "F") %>% count(wt=prop) #0.002612356

Arielnames <- filter(babynames, str_detect(babynames$name,"Ar[iy]+.l+[^a]?$")) %>%
  group_by(year) %>%
  summarize(proportionAriel = sum(prop))

Allnames <- filter(babynames, str_detect(babynames$name,"A")) %>% group_by(year) %>% summarize(proportionAlll = sum(prop))

ggplot(data = Arielnames) + geom_line(aes(x = year, y = proportionAriel), color = "green") + geom_vline(xintercept = 1989) + ggtitle("The popularity of the name 'Ariel' over time")

ggplot(data = Arielnames) + geom_line(aes(x = year, y = proportionAriel), color = "blue") + geom_vline(xintercept = 1989) + ggtitle("The popularity of the name 'Ariel' in 1988 and 1990") + coord_cartesian(xlim = c(1988,1990))

ggplot(data = Allnames) + geom_line(aes(x = year, y = proportionAlll), color = "red") + geom_vline(xintercept = 1989) + ggtitle("The popularity of all 'A' names over time") + coord_cartesian(xlim = c(2000,2017))
```
####The Change in proportion in female Ariel baby names
The popularity of the name 'Ariel' increased in popularity between 1988 and 1990. The movie was released in the 1989 between the year range where it increased drastically. In 1988, the proportion of popularity was higher than usual being 0.0008, but then jumped to 0.0026 in 1990 after the release of the movie. The percentile of change in Ariel names ends up being 1, which results to the 100th percentile. No other girl names had the a bigger difference in proportions than the name 'Ariel'. The proportion of names besides 'Ariel' were higher in comparison to babies with the name, as shown on the plots above. At the year 1989, the proportion of 'Ariel' names boomed to almost a proportion of 0.004, but then slowly declined after the year of the movie 'The little Mermaid' being released.Then, looking at the data of the exact year 1989, the proportion of the name 'Ariel' was on a rapid incline after the movie was released. In comparison to other names besides 'Ariel', the proportion trend remainds the same as the years go on, but a slight increase in the beginning of the early 2000s as the popularity of 'Ariel' was on a decline.

###Adam's Findings

```{r}
library(tidyverse)
library(babynames)



names <- babynames %>%
  group_by(year)%>%
  summarise(total = sum(n))

ada <- babynames %>%
  filter(str_detect(name, "^Ada"))%>%
  group_by(year)%>%
  summarise(num_ada = sum(n))

ada2 <- babynames %>%
  filter(str_detect(name, "^Adam$"))%>%
  group_by(year)%>%
  summarise(num_adam = sum(n))

newdata <- ada %>%
  left_join(names, by = "year")%>%
  mutate(prop = num_ada / total)

newnewdata <- newdata %>%
  left_join(ada2, by = "year") %>%
  mutate(prop_adam = num_adam / total)

ggplot(data = newnewdata)+
  geom_line(mapping = aes(x = year, y = prop), color = "red")+
  geom_line(mapping = aes(x = year, y = prop_adam))+
  labs(title = "Prop of names starting with Ada(red) and babies named Adam(black) by year")

FromBirth <- newnewdata %>%
  select(year,num_adam, prop_adam)%>%
  filter(between(year,1996,2017))
  

ggplot(data = FromBirth)+
  geom_line(mapping = aes(x = year, y = prop_adam))+
  labs(title = "Prop of babies named 'Adam' from 1996 - 2017")

BeforeBirth <- newnewdata %>%
  select(year, num_adam, prop_adam)%>%
  filter(between(year,1974,1996))

ggplot(data = BeforeBirth)+
  geom_line(mapping = aes(x = year, y = prop_adam))+
  labs(title = "Prop of babies named 'Adam' from 1974 - 1996")

```

####Comments
From my first graph I found that there were a much higher proportion of names that started with "Ada" than there were babies actually named "Adam". However this trend changed around 1960, and from about 1960 - 2000 almost all babies with names starting with "Ada" were actually named "Adam", this trend again changed around 2000 where the proportion of babies named "Adam" began to decline, while the proportion of babies with names starting with "Ada" began to increase rather sharply.  From my second set of graphs, I found that there was a spike in popularity of the name Adam from about 1975 - 1985, and then from then until 2017 there has been a steady decline in popularity of the name.


```{r}
percent_new <- newnewdata %>%
  mutate(percent = prop_adam * 100)

quantile(percent_new$percent)
```

####More comments
These percentiles demonstrate the percentage of babies named "Adam" relative to other baby names in each year.  The 50th percentile shows that on average, most years recorded around 3.1% of baby names to be "Adam".  The 100th percentile shows that on very rare occasions there were actually years that recorded close to 70% of baby names to be Adam.

### Richard's Findings

```{r}
library(tidyverse)
library(babynames)

#data <- babynames

NamesRIC <- babynames %>%
  filter(str_detect(name, "^Ric") == T)

PropName <- babynames %>%
  group_by(year) %>%
  summarise(number = sum(n))

PropRIC <- babynames %>%
  group_by(year) %>%
  filter(str_detect(name, "^Ric") == T) %>%
  summarise(number_RIC = sum(n))

PropRIC2 <- babynames %>%
  filter(str_detect(name, "^Richard$"))%>%
  group_by(year)%>%
  summarise(number_RICHARD = sum(n))


Newnamedata <- PropRIC %>%
  left_join(PropName, by = "year") %>%
  mutate(prop = number_RIC / number)

Newnamesdata <- Newnamedata %>%
  left_join(PropRIC2, by = "year") %>%
  mutate(proportion_Richard = number_RICHARD / number)

ggplot(data = Newnamesdata)+
  geom_point(mapping = aes(x = year, y = prop), color = "Blue")+
  geom_point(mapping = aes(x = year, y = proportion_Richard),color = "Red")+
  labs(title = "The Proportion of names that start with Ric(in blue) and babies named Richard(in red) by each year")

BirthYearName <- babynames %>%
  filter(year == 1999) %>%
  filter(str_detect(name, "^Richard$")) %>%
  summarise(number_Richardbirth = sum(n)) %>%
  mutate(number_Richardbirth, total1 = sum(number_Richardbirth))

CurrentName <- babynames %>%
  filter(year == 2017) %>%
  filter(str_detect(name, "^Richard$")) %>%
  summarise(number_Richardcurrent = sum(n))


BeforeBirthYearName <- babynames %>%
  filter(year == 1981) %>%
  filter(str_detect(name, "^Richard$")) %>%
  summarise(number_Richardbefore = sum(n)) %>%
  mutate(number_Richardbefore, total = sum(number_Richardbefore))

Richard_percent1 <- as_tibble(BeforeBirthYearName)%>%
  mutate(percentile = ((number_Richardbefore/total)/(1981-1999))*100)


Richard_percent2 <- as_tibble(BirthYearName) %>%
  mutate(percentile2 = ((number_Richardbirth/total1)/(1999-2017))*100)

ggplot(Richard_percent1, aes(number_Richardbefore), fill = Richardbefore) +
geom_histogram(binwidth = 1)

ggplot(Richard_percent2, aes(number_Richardbirth), fill = Richardbirth) +
geom_histogram(binwidth = 1)
```
####Comments
To begin, I had to find the proportions of names that began with "RIC" or my full name "Richard," so I utilized the left join function to combine the babynames dataset. My name's popularity compared to when I was born (in 1999) has declined. As seen in the ggplot, the red dots are on a downward slope which indicates that there was a downward decline in popularity. Furthermore, my name is currently not as popular compared to other names that start with the same 3 letters because the blue dots (which indicate the proportion of all names that start with "Ric") is higher than the red dots which shows that there are a lot of other names that start with "Ric". If the blue dots and red dots were a lot closer together then my name would be extremely popular among babies whose names started with "Ric". Lastly, my name was much more popular in my birth year (with 6514 babies named Richard) compared to 2017 which had only 2269 babies named Richard. That was a total difference of 4245 baby names! Further analysis revealed that my name 18 years before my birth year was, again, much more popular. In 1981, the number of babies named Richard amounted to 16972, whereas, my birth year only had 6514 babies named Richard. That was a difference of 10458 baby names, which also emphasizes how drastically unpopular my name is becoming.

###Erin's Findings
```{r}
library(tidyverse)
library(babynames)

dataset <- babynames

selectName <- babynames %>%
  group_by(year) %>%
  summarise(total = sum(n))

ERI_names <- babynames %>%
  filter(str_detect(name, "^Eri")) %>%
  group_by(year) %>%
  summarise(num_ERI = sum(n))

ERIN_names <- babynames %>%
  filter(str_detect(name, "^Erin$")) %>%
  group_by(year) %>%
  summarise(num_ERIN = sum(n))

data2 <- ERI_names %>%
  left_join(selectName, by = "year") %>%
  mutate(proportion = num_ERI / total)

data3 <- data2 %>%
  left_join(ERIN_names, by = "year") %>%
  mutate(proportionERIN = num_ERIN / total)

ggplot(data = data3) +
  geom_line(mapping = aes(x = year, y = proportion), color = "pink") + 
  geom_line(mapping = aes(x = year, y = proportionERIN), color = "blue") + 
  labs(title = "Proportion of names starting with 'ERI'(pink) compared to the name 'ERIN'(blue) over time")

BirthYearErin <- data3 %>%
  select(year,num_ERIN, proportionERIN)%>%
  filter(between(year,2000,2017))

ggplot(data = BirthYearErin)+
  geom_line(mapping = aes(x = year, y = proportionERIN))+
  labs(title = "Proportion of babies named 'Erin' from 2000 to 2017")

BeforeBirthERIN <- data3 %>%
  select(year, num_ERIN, proportionERIN)%>%
  filter(between(year,1983,2000))

ggplot(data = BeforeBirthERIN)+
  geom_line(mapping = aes(x = year, y = proportionERIN))+
  labs(title = "Proportion of babies named 'Erin' from 1983 to 2000")
```
####Comments
In order to find the proportions of names that began with "ERI" or my full name "ERIN", I incorporated the join function to join the babynames dataset for the different manipulated names. As shown by the data, my name wasn't popular until the 1980s where names beginning with "Eri" and named "Erin" peaked. Then, around the time I was born the popularity was declining. The popularity of a name starting with "Eri" was higher than "Erin", because there are various manipulations of name that involve the same three letters, which causes the higher proportion in pink on the graph. 

###Jasmine's Findings
```{r}
babies <- babynames

names <- babies %>%
  group_by(year)%>%
  summarise(total = sum(n))

jas <- babies %>%
  filter(str_detect(name, "^Jas"))%>%
  group_by(year)%>%
  summarise(num_jas = sum(n))

jas2 <- babies %>%
  filter(str_detect(name, "^Jasmine$"))%>%
  group_by(year)%>%
  summarise(num_Jasmine = sum(n))

new_babies <- jas %>%
  left_join(names, by = "year")%>%
  mutate(num_jas, jas_prop = num_jas / total)


new_babies2 <- new_babies %>%
  left_join(jas2, by = "year") %>%
  mutate(num_Jasmine, Jasmine_prop = num_Jasmine / total)



ggplot(data = new_babies2)+
  geom_col(mapping = aes(x = year, y = Jasmine_prop), color = 'green', na.rm = TRUE, position = "stack")+
   labs(title = "Prop of babies named Jasmine")
ggplot(data = new_babies2)+
  geom_col(mapping = aes(x = year, y = jas_prop), color = 'purple', na.rm = TRUE, position = "stack")+
  labs(title = "Prop of names starting with Jas")
```

####Initial Comments

```{r}
jasmine1 <- babies %>%
  filter(str_detect(name, "^Jasmine$"))%>%
  filter(year > 1998)%>%
  group_by(year)%>%
  summarise(num_Jasmine1 = sum(n))%>%
  mutate(num_Jasmine1, total = sum(num_Jasmine1))

jasmine2 <- babies %>%
  filter(str_detect(name, "^Jasmine$"))%>%
  filter(year < 2000, year > 1980)%>%
  group_by(year)%>%
  summarise(num_Jasmine2 = sum(n))%>%
  mutate(num_Jasmine2, total1 = sum(num_Jasmine2))

J_percentile1 <- as.tibble(jasmine1)%>%
  mutate(percentile = ((num_Jasmine1/total)/(1999-1981))*100)
J_percentile1

J_percentile2 <- as.tibble(jasmine2) %>%
  mutate(percentile1 = ((num_Jasmine2/total1)/(2017- 1999))*100)
J_percentile2
  

```
####Comments Continued 


```{r}
#ggplot(J_percentile1, aes(num_Jasmine1), fill = jasmine1) +
#  geom_histogram(binwidth = 0.1)
```
#### Final Comments

###WHO DID WHAT
Adam created new data tables and calculated proportions of babies named Adam compared to babies with names starting with "Ada" and graphed the two together.  He also created graphs showing the trend in popularity of his name over the 22 years before he was born, until 2017.  He then calculated the percentiles of percentage of babies named Adam by year. Richard created data tables and sets based on babies named Richard or babies that started with "RIC." Richard then used those new data tables to compare the popularity of "Richard's" among names that started with "RIC." As mentioned before, graphs were created to depict the trend in popularity of babies named Richard 18 years before his birth year, at his birth year (which was 1999), and in 2017. Percentages indicated that Richard was on a downward decline in terms of popularity and Richard's name was much more popular at his birth year compared to 2017. Erin contributed to the two group tasks of finding the Rachel and Ariel numbers and name differences. She was able to do this by filtering for the babynames$names to the specifics, then she used str_detect to count for the string variation names of Ariel and Rachel. Overall, I discovered that Rachel is a more common name than Ariel has been over the years. Then, for The Little Mermaid, I was able to do similar actions for the previous lab question, but I used proportions along with str_detect. For Erin's individual part, she contributed her data for 'Erin' and found it was popular around the 1980s and has been a slow decline ever since. 